---
title: "ACTIVITY 1"
author: "David Tofan"
date: "18/5/2020"
output:
  html_document:
    theme: readable
    highlight: tango
    toc: true
    toc_depth: 2
    toc_float: true
    toc_collapsed: true
    number_sections: true
---

# Description

This was an activity during our university class Geolocation.

# Load libraries

```{r}
# Install/Load all necessary libraries
prepare_packages <- function(packages){
  # Check what packages are not installed
  not_installed <- packages[!(packages %in% installed.packages()[, "Package"])]
  # Install missing packages
  if (length(not_installed)) 
    install.packages(not_installed, dependencies = TRUE)
  # Load the entire list of packages
  sapply(packages, require, character.only = TRUE)
}
packages <- c("rgdal",
              "sp",
              "raster",
              "rgeos",
              'MASS'
)
prepare_packages(packages)
```

# Question 1

Download a Shape File.

* How many Polygons does it have?

Dimensions: 159, 20

```{r}
# SOURCE: https://arc-garc.opendata.arcgis.com/datasets/dc20713282734a73abe990995de40497_68
# OGA Open User Licence
# Georgia
# Load data (Polygons)
Poligonos_Georgia <- readOGR(dsn="/Actividades/Counties_Georgia-shp/Counties_Georgia.shp", layer="Counties_Georgia")

# ANOTHER METHOD TO DOWNLOAD DATA
#library(geosphere)
# Download shapefile of Georgia’s borders
#Poligonos_Georgia = getData("GADM", country="GEO", level=0)
```

```{r}
# Visualize
par(mfrow = c(1, 2))
plot(Poligonos_Georgia, main="Georgia")
```

```{r}
# Dimensions
dim(Poligonos_Georgia)
```

```{r}
# CRS
#crs(Poligonos_Georgia)
Poligonos_Georgia@proj4string

# Change CRS
Poligonos_Georgia_WGS84 <- spTransform(Poligonos_Georgia, CRS("+proj=longlat +datum=WGS84"))
plot(Poligonos_Georgia_WGS84, main = "Poligonos_WGS84")
```

```{r}
# Convert to DataFrame
DF_Poligonos_Georgia <- as.data.frame(Poligonos_Georgia)
head(DF_Poligonos_Georgia)
```

# Question 2

* After reading the SHP File, what class does the created object have? (Class (object))

```{r}
class(Poligonos_Georgia)
```

```{r}
class(DF_Poligonos_Georgia)
```

# Question 3

* Which are the 3 regions with the largest area?

Ware, Burke, and Clinch County have the biggest areas.

```{r}
# Calculate area (m2)
sqm <- geosphere::areaPolygon(Poligonos_Georgia)
# Convert sqm to km2
sqkm <- sqm/1000000
# Add as column
Poligonos_Georgia$Area_sqm <- sqkm
DF_Poligonos_Georgia$Area_sqm <- sqkm

# Area, add as column (?)
#Poligonos_Georgia$Area_sqm <- area(Poligonos_Georgia)
```

```{r}
# Exctract the 3 biggest areas
Pol_ordenados <- order(Poligonos_Georgia$Area_sqm,
                       decreasing = TRUE)[1:3]
as.data.frame(Poligonos_Georgia[Pol_ordenados,])
```

# Question 4

* Download 5 coordinates of specific points of the region through any official source.

Random clicks on the map (show address) in the State of Georgia.
Source: https://www.openstreetmap.org/search

```{r}
par(mfrow = c(1, 2))
Latitud <- c(32.9718,33.7083,34.0140,32.6301,32.389)
Longitud <- c(-84.2651,-84.4135,-84.3558,-83.6169,-81.914)
LONGLAT <- as.data.frame(cbind(Longitud, Latitud))
LONGLAT
```

```{r}
bbdd2 <- LONGLAT
# Create coordinates
coordinates(bbdd2) <- c('Longitud','Latitud')
# Add WGS84 as CRS
proj4string(bbdd2) <- proj4string(Poligonos_Georgia)
```

# Question 5

* Represent the Polygons and the Points together.

<b>First Method</b>

```{r}
plot(Poligonos_Georgia, main="Poligonos")
points(bbdd2, cex=2, pch=19, col="red")
```

<b>Segundo método</b>

```{r}
Poligonos_WGS84 <- spTransform(Poligonos_Georgia,CRS("+proj=longlat +datum=WGS84"))
plot(Poligonos_WGS84,main="Poligonos_WGS84")

points(bbdd2,cex=1,pch=19,col="red")
```

# Question 6

* Could you briefly describe the following coordinate system?

`coord. ref. : +proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0`

CRS: Coordinate Reference System

`+proj=longlat`: compact way to identify a spatial or coordinate reference system. In this case it is `longlat`, meaning Longitude and Latitude.

`+datum=WGS84`: The datum WGS84 (the datum refers to the 0,0 reference for the coordinate system used in the projection).

`+ellps=WGS84`: the ellipsoid (how the earth’s roundness is calculated) for the data is WGS84.

`+towgs84=0,0,0`: it is an additional that shows us the preferred transformation for our `datum`.


